---
title: "Beetle Kill"
author: "Lauren"
date: "8/14/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pole Creek Fire Data

The purpose of this study was to see if the amount / chemical structure of PyC that is produced during a forest fire depends on whether the trees were alive at the time of fire.

-	Two samples (1. scorched bark and 2. charred wood from the base of the tree) were collected from each of three trees that had been killed by beetles prior to the Pole Creek Fire (PCF). Beetle kill was confirmed by evidence of J-shaped galleries under the bark. 
-	Two samples (1. Scorched bark and 2. Wood from under scorched bark @ base of tree) were collected from each of three trees that had been alive at the time of the fire. 

Samples were dried O/N in oven at 50C on 6/20/18

PCF samples were ground in a ball grinder on 6/21/18
- NOTE: the scorched bark from Tree 4 and Tree 5 (live @ time of fire) still contained a good bit of sap residue. Ball grinding resulted in a tacky mashed up thing. Transfer to scin vial and dry @ 50C for an additional 3 days. Still seemed tacky and heterogeneous. Flash froze in liquid nitrogen and ground in mortar and pestle to a fine powder. 

***


#Metadata

* jh.number:	index number used for internal tracking
* name:	sample name used during analysis
* field.fire:	was the sample collected from the pole creek fire (PCF) or from the trees anna burned in the lab (lab)
* tree:	individual tree number combined with the pcf/lab designation
* base:	two of the samples anna burned in the lab were designated as 'base' - I'm not exactly sure what that means
* live.dead:	was the tree alive or dead at the time of the fire / burning
* bark.char:	was the sample from the bark or exposed char/wood under burned bark
* bpca.carbon:	g total bpca / kg carbon
* bpca.soil:	g total bpca / kg material
* b6.total:	b6ca:total bpca ratio

***

# Packages

```{r warning=FALSE,message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(lme4)
library(lsmeans)
library(statmod)
library(car)
library(DHARMa)
library(GGally)
library(MASS)
library(cowplot)
library(grid)
library(gtable)
```

# Data
```{r}
bk=read.csv("../data/beetle_kill.csv", header=TRUE,sep=",",na.strings=c("", "NA"))
```

## Pole Creak data
```{r}
pcf=subset(bk,field.fire=="PCF")
```

Add column for beetle-killed and fire-killed category
ord_class = c("fire", "beetle")
 %>% 
  mutate(kill_by = factor(kill_by, levels = ord_class))
```{r}
pcf1 = pcf %>% 
  mutate(kill_by = ifelse(live.dead=="dead", "beetle", "fire")) 
pcf1
```

# Data Visualization

```{r}
qplot(x = kill_by, y = bpca.carbon, data = pcf1, geom = "boxplot") +
stat_summary(fun = mean, aes(shape = "mean"), geom = "point", color = "red", size = 3)
```

```{r}
qplot(x = kill_by, y = bpca.soil, data = pcf1, geom = "boxplot") +
stat_summary(fun = mean, aes(shape = "mean"), geom = "point", color = "red", size = 3)
```
```{r}
qplot(x = kill_by, y = b6.total, data = pcf1, geom = "boxplot") +
stat_summary(fun = mean, aes(shape = "mean"), geom = "point", color = "red", size = 3)
```

# Analysis

## With Linear Model (lm)
### BPCA carbon
#### Summary data
```{r}
( sumdat1 = summarise( group_by(pcf1, "Mortality agent" = kill_by),
n = n(),
Mean = mean(bpca.carbon),
SD = sd(bpca.carbon),
Median = median(bpca.carbon),
Minimum = min(bpca.carbon),
Maximum = max(bpca.carbon) ) )
```
#### Model
```{r}
fit1 = lm(bpca.carbon ~ kill_by, data = pcf1)
```

#### Residuals
```{r}
# Get the fitted value for each observation
pcf1$fit1 = fitted(fit1)
# Get the residuals of the model
pcf1$res1 = residuals(fit1)
```

```{r}
qplot(x = fit1, y = res1, data = pcf1,
main = "Residuals vs Fitted Values")
```

```{r}
qplot(x = kill_by, y = res1, data = pcf1,
xlab = "Fire-killed v Beetle-killed group",
ylab = "Residuals",
main = "Residuals vs Mortality group")
```
```{r}
qplot(x = factor(1), y = res1, data = pcf1, geom = "boxplot")
```

```{r}
qqnorm(pcf1$res1, main = "Normal QQ Plot of Residuals")
qqline(pcf1$res1) # add reference line to the qq plot
```

```{r}
plot(fit1, which = 1) # residual vs fitted values
```

```{r}
plot(fit1, which = 2) # qqnorm plot of residuals
```

#### Summary
```{r}
summary(fit1)
```
```{r}
# All the coefficients
coef(fit1)
```

```{r}
# All the confidence intervals for the coefficients
confint(fit1)
```

#### t-test
```{r}
( respunequal = t.test(bpca.carbon ~ kill_by, data = pcf1) )
```

##### Table
"g BPCA /\nkg material"
```{r}
# Write out the results from the two-sample t-test with unequal variances
# A title for the test
titleres = as.character(expression(bold("Welch"~t-test*~"for"~a~difference*~"in"~means)))
# This is the t stat with df and p-value
ttestres1 = as.character(expression(t[6.5]==-4.7*","~italic(p)==0.003))
# The estimated difference in group means (I calculated from the test results)
diff1 = as.character(expression(Delta==21.8~g~BPCA~kg~C^-1))
# The 95% CI of the difference
ci1 = as.character(expression(95*"%"~CI*":"~10.6~to~33.0))
# Put these into a single column of a data.frame
d1 = data.frame(col = rbind(ttestres1, diff1, ci1) )
```

```{r}
t1 = tableGrob(d1, rows = NULL, cols = NULL,
theme = ttheme_default(core = list(bg_params = list(fill = "white"),
fg_params = list(parse = TRUE),
padding = unit(c(2.5, 2.5), "mm"))))
```

```{r}
t1 = gtable_add_grob(t1, grobs = rectGrob(gp = gpar(fill = NA, lwd = NA)),
t = 1, b = nrow(t1), r = ncol(t1), l = 1)
```

#might not need
```{r}
( diff_fit1 = data.frame(coef(fit1),
confint(fit1)) )
```
```{r}
# Rename column names
names(diff_fit1) = c("Estimate", "Lower", "Upper")
diff_fit1
```
```{r}
rownames(diff_fit1)
```


```{r}
( diff_fit1$kill_by = factor(rownames(diff_fit1),
levels = c("(Intercept)",   "kill_bybeetle"),
labels = c("fire", "beetle") ) )
```

#### Graph

```{r}
ggplot() + 
      geom_point(data = diff_fit1, aes(x = kill_by, y = Estimate, color = kill_by), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = diff_fit1, width = .2, lwd = .75, aes(x = kill_by, y = Estimate, ymin = Lower, ymax = Upper, color = kill_by), position = position_dodge(width = .75)) + 
 
    geom_point(data= pcf1, aes(x = kill_by, y = bpca.carbon, color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg C") +
  scale_color_manual(values = c("darkgreen", "tan")) + 
  
scale_x_discrete(labels= c("Fire-killed", "Beetle-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

```
#### Boxplot
```{r}
g1 = ggplot(pcf1, aes(x = kill_by, y = bpca.carbon, color=kill_by) ) + # define plot axes
  geom_boxplot(outlier.shape = NA) + # Add boxplots for each group
  geom_point(aes(color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
stat_summary(fun = mean, geom= "point", size = 4) + # Add means to the plot
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg C") +
  
  scale_color_manual(values = c("tan", "darkgreen")) +
  scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))


g1
```

### BPCA Mass
#### Summary data
```{r}
( sumdat2 = summarise( group_by(pcf1, "Mortality agent" = kill_by),
n = n(),
Mean = mean(bpca.soil),
SD = sd(bpca.soil),
Median = median(bpca.soil),
Minimum = min(bpca.soil),
Maximum = max(bpca.soil) ) )
```
#### Model
```{r}
fit2 = lm(bpca.soil ~ kill_by, data = pcf1)
```
```{r}
head(pcf1)
```

#### Residuals
```{r}
# Get the fitted value for each observation
pcf1$fit2 = fitted(fit2)
# Get the residuals of the model
pcf1$res2 = residuals(fit2)
```

```{r}
qplot(x = fit2, y = res2, data = pcf1,
main = "Residuals vs Fitted Values")
```

```{r}
qplot(x = kill_by, y = res2, data = pcf1,
xlab = "Fire-killed v Beetle-killed group",
ylab = "Residuals",
main = "Residuals vs Mortality group")
```
```{r}
qplot(x = factor(1), y = res2, data = pcf1, geom = "boxplot")
```

```{r}
qqnorm(pcf1$res2, main = "Normal QQ Plot of Residuals")
qqline(pcf1$res2) # add reference line to the qq plot
```

```{r}
plot(fit2, which = 1) # residual vs fitted values
```

```{r}
plot(fit2, which = 2) # qqnorm plot of residuals
```

#### Summary
```{r}
summary(fit2)
```
```{r}
# All the coefficients
coef(fit2)
```

```{r}
# All the confidence intervals for the coefficients
confint(fit2)
```

#### t-test
```{r}
( fit2_respunequal = t.test(bpca.soil ~ kill_by, data = pcf1) )
```

##### Table
"g BPCA /\nkg material"
```{r}
# Write out the results from the two-sample t-test with unequal variances
# A title for the test
titleres = as.character(expression(bold("Welch"~t-test*~"for"~a~difference*~"in"~means)))
# This is the t stat with df and p-value
ttestres2 = as.character(expression(t[5.8]==-4.2*","~italic(p)==0.006))
# The estimated difference in group means (I calculated from the test results)
diff2 = as.character(expression(Delta==15.0~g~BPCA~kg~material^-1))
# The 95% CI of the difference
ci2 = as.character(expression(95*"%"~CI*":"~6.2~to~23.8))
# Put these into a single column of a data.frame
d2 = data.frame(col = rbind(ttestres2, diff2, ci2) )
```

```{r}
t2 = tableGrob(d2, rows = NULL, cols = NULL,
theme = ttheme_default(core = list(bg_params = list(fill = "white"),
fg_params = list(parse = TRUE),
padding = unit(c(2.5, 2.5), "mm"))))
```

```{r}
t2 = gtable_add_grob(t2, grobs = rectGrob(gp = gpar(fill = NA, lwd = NA)),
t = 1, b = nrow(t2), r = ncol(t2), l = 1)
```

#### Model estimates (delete?)
```{r}
( diff_fit2 = data.frame(coef(fit2),
confint(fit2)) )
```
```{r}
# Rename column names
names(diff_fit2) = c("Estimate", "Lower", "Upper")
diff_fit2
```
```{r}
rownames(diff_fit2)
```


```{r}
( diff_fit2$kill_by = factor(rownames(diff_fit2),
levels = c("(Intercept)",   "kill_bybeetle"),
labels = c("fire", "beetle") ) )
```

#### Graph

```{r}
ggplot() + 
      geom_point(data = diff_fit2, aes(x = kill_by, y = Estimate, color = kill_by), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = diff_fit2, width = .2, lwd = .75, aes(x = kill_by, y = Estimate, ymin = Lower, ymax = Upper, color = kill_by), position = position_dodge(width = .75)) + 
 
    geom_point(data= pcf1, aes(x = kill_by, y = bpca.soil, color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg material") +
  scale_color_manual(values = c("darkgreen", "tan")) + 
  
scale_x_discrete(labels= c("Fire-killed", "Beetle-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

```
#### Boxplot
```{r}
g2 = ggplot(pcf1, aes(x = kill_by, y = bpca.soil, color=kill_by) ) + # define plot axes
  geom_boxplot(outlier.shape = NA) + # Add boxplots for each group
  geom_point(aes(color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
stat_summary(fun = mean, geom= "point", size = 4) + # Add means to the plot
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg material") +
  
  scale_color_manual(values = c("tan", "darkgreen")) +
  scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))


g2
```

### B6CA Total

#### Summary data

```{r}
( sumdat3 = summarise( group_by(pcf1, "Mortality agent" = kill_by),
n = n(),
Mean = mean(b6.total),
SD = sd(b6.total),
Median = median(b6.total),
Minimum = min(b6.total),
Maximum = max(b6.total) ) )
```
#### Model
```{r}
fit3 = lm(b6.total ~ kill_by, data = pcf1)
```

#### Residuals
```{r}
# Get the fitted value for each observation
pcf1$fit3 = fitted(fit3)
# Get the residuals of the model
pcf1$res3 = residuals(fit3)
```

```{r}
qplot(x = fit3, y = res3, data = pcf1,
main = "Residuals vs Fitted Values")
```

```{r}
qplot(x = kill_by, y = res3, data = pcf1,
xlab = "Fire-killed v Beetle-killed group",
ylab = "Residuals",
main = "Residuals vs Mortality group")
```
```{r}
qplot(x = factor(1), y = res3, data = pcf1, geom = "boxplot")
```

```{r}
qqnorm(pcf1$res3, main = "Normal QQ Plot of Residuals")
qqline(pcf1$res3) # add reference line to the qq plot
```

```{r}
plot(fit3, which = 1) # residual vs fitted values
```

```{r}
plot(fit3, which = 2) # qqnorm plot of residuals
```

#### Summary
```{r}
summary(fit3)
```
```{r}
# All the coefficients
coef(fit3)
```

```{r}
# All the confidence intervals for the coefficients
confint(fit3)
```

#### t-test
```{r}
( respunequal = t.test(b6.total ~ kill_by, data = pcf1) )
```


##### Table
Difference in means
```{r}
0.2116667-0.2233333
```

g B6CA /\nTotal BPCA
```{r}
# Write out the results from the two-sample t-test with unequal variances
# A title for the test
titleres = as.character(expression(bold("Welch"~t-test*~"for"~a~difference*~"in"~means)))
# This is the t stat with df and p-value
ttestres3 = as.character(expression(t[5.09]==-0.18*","~italic(p)==0.86))
# The estimated difference in group means (I calculated from the test results)
diff3 = as.character(expression(Delta==-0.012~g~B6CA~Total~BPCA^-1))
# The 95% CI of the difference
ci3 = as.character(expression(95*"%"~CI*":"~-0.17~to~0.15))
# Put these into a single column of a data.frame
d3 = data.frame(col = rbind(ttestres3, diff3, ci3) )
```

```{r}
t3 = tableGrob(d3, rows = NULL, cols = NULL,
theme = ttheme_default(core = list(bg_params = list(fill = "white"),
fg_params = list(parse = TRUE),
padding = unit(c(2.5, 2.5), "mm"))))
```

```{r}
t3 = gtable_add_grob(t3, grobs = rectGrob(gp = gpar(fill = NA, lwd = NA)),
t = 1, b = nrow(t3), r = ncol(t3), l = 1)
```

#### Model estimates (delete?)

```{r}
( diff_fit3 = data.frame(coef(fit3),
confint(fit3)) )
```

```{r}
# Rename column names
names(diff_fit3) = c("Estimate", "Lower", "Upper")
diff_fit3
```
```{r}
rownames(diff_fit3)
```


```{r}
( diff_fit3$kill_by = factor(rownames(diff_fit3),
levels = c("(Intercept)",   "kill_bybeetle"),
labels = c("fire", "beetle") ) )
```

#### Graph

```{r}
ggplot() + 
      geom_point(data = diff_fit3, aes(x = kill_by, y = Estimate, color = kill_by), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = diff_fit3, width = .2, lwd = .75, aes(x = kill_by, y = Estimate, ymin = Lower, ymax = Upper, color = kill_by), position = position_dodge(width = .75)) + 
 
    geom_point(data= pcf1, aes(x = kill_by, y = b6.total, color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_color_manual(values = c("darkgreen", "tan")) + 
  
scale_x_discrete(labels= c("Fire-killed", "Beetle-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

```
#### Boxplot
```{r}
g3 = ggplot(pcf1, aes(x = kill_by, y = b6.total, color=kill_by) ) + # define plot axes
  geom_boxplot(outlier.shape = NA) + # Add boxplots for each group
  geom_point(aes(color = kill_by, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
stat_summary(fun = mean, geom= "point", size = 4) + # Add means to the plot
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g B6CA /\nTotal BPCA") +
  
  scale_color_manual(values = c("tan", "darkgreen")) +
  
scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))


g3
```

## Manuscript figure

### Legend

```{r}
legend_bottom = ggplot(pcf, aes(x = live.dead, y = b6.total)) +
  geom_jitter(aes(shape = bark.char), size = 3) +
  labs(title = "ACI of PyC", x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_shape_manual(values = c(23, 24)) +
  theme_bw() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
              legend.text = element_text(size = 9)) +  
  theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
theme(panel.grid.major = element_blank())+
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

legend_bottom
  
```

```{r}
legend_b = get_legend(legend_bottom)
```


### Gridded boxplots

```{r fig.height=7, fig.width=6}

p1 = cowplot::plot_grid(g1, g2, g3, labels = c("A", "B", "C"), ncol = 1) 
p1
```

### Gridded with t-test data

```{r fig.height=7, fig.width=6}
p2 = p1 + 
  draw_grob(t1,x =.16, y = .80, width = .25, height = .25, scale=.5) +
  draw_grob(t2,x =.185, y = .465, width = .25, height = .25, scale=.5) +
  draw_grob(t3,x =.653, y = .133, width = .25, height = .25, scale=.5)

p2
```

### Gridded with t-test data leg bot
```{r fig.height=7, fig.width=6}


p3 = plot_grid(p2, legend_b, ncol = 1, rel_heights = c(1, .1))

p3
```

### Gridded with t-test data

```{r fig.height=7, fig.width=6}
p4 = p1 + 
  draw_plot(legend_b, x =.10, y = .59, width = .3, height = .3) + 
  draw_grob(t1,x =.66, y = .80, width = .25, height = .25, scale=.5) +
  draw_grob(t2,x =.66, y = .465, width = .25, height = .25, scale=.5) +
  draw_grob(t3,x =.21, y = .135, width = .25, height = .25, scale=.5)

p4
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-10_PyC_live-dead_t-test.jpeg", plot = p4, width = 6, height = 7, units = c("in"), dpi=600 )
```

# GLM
## Model 1 - BPCA per Carbon
```{r}
mod1 = glm(bpca.carbon ~ live.dead, data = pcf, family = gaussian)
```



### Residuals with DHARMa
```{r, include=FALSE}
mod1.res = simulateResiduals(fittedModel = mod1, n = 500)
mod1.res$scaledResiduals
mod1.res$scaledResidualsNormal
```

In the plot below we are looking for the points in the QQ plot to fall along the RED linen. In the Residuals vs.predicted plot were are looking for the lines to be relatively horizontal. All looks good. 
```{r}
plotSimulatedResiduals(simulationOutput = mod1.res)
```

```{r}
testDispersion(mod1.res)
```

### Sumamry
```{r}
summary(mod1)
```

```{r}
summary(lsmeans(mod1, pairwise ~ live.dead), type = "response", infer = TRUE, level = .90, adjust = "bon")
```

### Data for Graph

```{r}
mod1_graph = data.frame(summary(lsmeans(mod1, pairwise ~ live.dead),type = "response"))
names(mod1_graph)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod1_graph
```

```{r eval=FALSE, include=FALSE}
levels(mod1_graph$lsmeans.live.dead)[levels(mod1_graph$lsmeans.live.dead)=="dead"] = "Charred"
levels(mod1_graph$lsmeans.live.dead)[levels(mod1_graph$lsmeans.live.dead)=="live"] = "Scorched"

( mod1_graph1 = dplyr::select(mod1_graph, lsmeans.live.dead:lsmeans.asymp.UCL))



names(mod1_graph1)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod1_graph1 = data.frame(mod1_graph1)


```

### Graph

```{r}
mod1_plot = ggplot(data = mod1_graph, aes(x = live.dead, y = response, color = live.dead) ) + 
   geom_errorbar(width = .2, lwd = .75, aes(ymin = lcl, ymax = ucl), position = position_dodge(width = .75)) + 
geom_point(size = 3, position = position_dodge(width = .75)) + 
  labs(x=(NULL), y= "g BPCA /\nkg C") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

mod1_plot  
```



### Graph with raw
```{r}
plot1_raw = ggplot() + 
      geom_point(data = mod1_graph, aes(x = live.dead, y = response, color = live.dead), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = mod1_graph, width = .2, lwd = .75, aes(x = live.dead, y = response, ymin = lcl, ymax = ucl, color = live.dead), position = position_dodge(width = .75)) + 
 
    geom_point(data= pcf, aes(x = live.dead, y = bpca.carbon, color = live.dead, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg C") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

plot1_raw
```
## Model 2 - BPCA per Material
```{r}
mod2 = glm(bpca.soil ~ live.dead, data = pcf, family = gaussian)
```

### Residuals with DHARMa
```{r, include=FALSE}
mod2.res = simulateResiduals(fittedModel = mod2, n = 250)
mod2.res$scaledResiduals
mod2.res$scaledResidualsNormal
```

In the plot below we are looking for the points in the QQ plot to fall along the RED linen. In the Residuals vs.predicted plot were are looking for the lines to be relatively horizontal. All looks good. 
```{r}
plotSimulatedResiduals(simulationOutput = mod2.res)
```

```{r}
testDispersion(mod2.res)
```

### Sumamry
```{r}
summary(mod2)
```

```{r}
summary(lsmeans(mod2, pairwise ~ live.dead), type = "response", infer = TRUE, level = .90, adjust = "bon")
```



### Data for Graph
```{r}
mod2_graph = data.frame(summary(lsmeans(mod2, pairwise ~ live.dead),type = "response"))
names(mod2_graph)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod2_graph
```

```{r eval=FALSE, include=FALSE}
levels(mod2_graph$lsmeans.live.dead)[levels(mod2_graph$lsmeans.live.dead)=="dead"] = "Charred"
levels(mod2_graph$lsmeans.live.dead)[levels(mod2_graph$lsmeans.live.dead)=="live"] = "Scorched"

( mod2_graph2 = dplyr::select(mod2_graph, lsmeans.live.dead:lsmeans.asymp.UCL))



names(mod2_graph2)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod2_graph2 = data.frame(mod2_graph2)


```


### Graph
```{r}
mod2_plot = ggplot(data = mod2_graph, aes(x = live.dead, y = response, color = live.dead) ) + 
   geom_errorbar(width = .2, lwd = .75, aes(ymin = lcl, ymax = ucl), position = position_dodge(width = .75)) + 
geom_point(size = 3, position = position_dodge(width = .75)) + 
  labs(x=(NULL), y= "g BPCA /\nkg material") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

mod2_plot  
```

### Graph with raw


```{r}
plot2_raw = ggplot() + 
      geom_point(data = mod2_graph, aes(x = live.dead, y = response, color = live.dead), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = mod2_graph, width = .2, lwd = .75, aes(x = live.dead, y = response, ymin = lcl, ymax = ucl, color = live.dead), position = position_dodge(width = .75)) + 
  scale_color_manual(values = c("tan", "darkgreen")) +
 
    geom_point(data= pcf, aes(x = live.dead, y = bpca.soil, color = live.dead, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g BPCA /\nkg material") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

plot2_raw
```

## Model 3 - B6 total
```{r}
mod3 = glm(b6.total ~ live.dead, data = pcf, family = gaussian)
```

### Residuals with DHARMa
```{r, include=FALSE}
mod3.res = simulateResiduals(fittedModel = mod3, n = 250)
mod3.res$scaledResiduals
mod3.res$scaledResidualsNormal
```

In the plot below we are looking for the points in the QQ plot to fall along the RED linen. In the Residuals vs.predicted plot were are looking for the lines to be relatively horizontal. All looks good. 
```{r}
plotSimulatedResiduals(simulationOutput = mod3.res)
```

```{r}
testDispersion(mod3.res)
```

### Sumamry
```{r}
summary(mod3)
```

```{r}
summary(lsmeans(mod3, pairwise ~ live.dead),type = "response", infer = TRUE, level = .9, adjust = "bon")
```


### Data for Graph
```{r}
mod3_graph = data.frame(summary(lsmeans(mod3, pairwise ~ live.dead),type = "response"))

names(mod3_graph)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod3_graph
```

```{r eval=FALSE, include=FALSE}
levels(mod3_graph$lsmeans.live.dead)[levels(mod3_graph$lsmeans.live.dead)=="dead"] = "Charred"
levels(mod3_graph$lsmeans.live.dead)[levels(mod3_graph$lsmeans.live.dead)=="live"] = "Scorched"

( mod3_graph3 = dplyr::select(mod3_graph, lsmeans.live.dead:lsmeans.asymp.UCL))



names(mod3_graph3)[1:6] = c("live.dead", "response", "se", "df", "lcl", "ucl")
mod3_graph3 = data.frame(mod3_graph3)


```


### Graph 
```{r}
mod3_plot = ggplot(data = mod3_graph, aes(x = live.dead, y = response, color = live.dead) ) + 
   geom_errorbar(width = .2, lwd = .75, aes(ymin = lcl, ymax = ucl), position = position_dodge(width = .75)) + 
geom_point(size = 3, position = position_dodge(width = .75)) + 
  labs(x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

mod3_plot  
```

### Graph with raw

```{r}
plot3_raw = ggplot() + 
      geom_point(data = mod3_graph, aes(x = live.dead, y = response, color = live.dead), size = 3, position = position_dodge(width = .75)) +  
      geom_errorbar(data = mod3_graph, width = .2, lwd = .75, aes(x = live.dead, y = response, ymin = lcl, ymax = ucl, color = live.dead), position = position_dodge(width = .75)) + 
  scale_color_manual(values = c("tan", "darkgreen")) +
 
    geom_point(data= pcf, aes(x = live.dead, y = b6.total, color = live.dead, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  scale_shape_manual(values = c(23, 24)) +
  labs(x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  scale_x_discrete(labels= c("Beetle-killed", "Fire-killed")) +
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

plot3_raw
```

## Manuscript Figures

### Without raw data
```{r fig.height=5.25, fig.width=4.75}

exp1_figure1 = cowplot::plot_grid(mod1_plot, mod2_plot, mod3_plot, labels = c("A", "B", "C"), ncol = 1) 
exp1_figure1
```

```{r fig.height=5.25, fig.width=4.75}
exp1_figure2 = exp1_figure1 + 
  draw_label("*", x = .55, y = .95, size = 36, fontface = "bold") + 
  draw_label("*", x = .55, y = .6, size = 36, fontface = "bold") 

exp1_figure2
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-07_exp1_dead-live_analysis.jpeg", plot = exp1_figure2, width = 4.75, height = 5.25, units = c("in"), dpi=600 )
```

### With Raw data

```{r}
legend_bottom = ggplot(pcf, aes(x = live.dead, y = b6.total)) +
  geom_jitter(aes(shape = bark.char), size = 3) +
  labs(title = "ACI of PyC", x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_shape_manual(values = c(23, 24)) +
  theme_bw() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
              legend.text = element_text(size = 9)) +  
  theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
theme(panel.grid.major = element_blank())+
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

legend_bottom
  
```



```{r fig.height=5.25, fig.width=4.75}

grid_fig2 = cowplot::plot_grid(plot1_raw, plot2_raw, plot3_raw, labels = c("A", "B", "C"), ncol = 1) 
grid_fig2
```

```{r fig.height=5.25, fig.width=4.75}
figure2 = exp1_figure1 + 
  draw_label("*", x = .55, y = .95, size = 36, fontface = "bold") + 
  draw_label("*", x = .55, y = .6, size = 36, fontface = "bold") 

exp1_figure2
```



```{r fig.height=5.25, fig.width=4.75}
legend2 = get_legend(legend_bottom)
grid_fig3 = ggdraw(grid_fig2) +
  draw_plot(legend2, x =.65, y = .78, width = .3, height = .3) + 
  draw_label("*", x = .55, y = .94, size = 36, fontface = "bold") + 
  draw_label("*", x = .55, y = .6, size = 36, fontface = "bold") 

grid_fig3

```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2020-07-07_PyC_live-dead_raw.jpeg", plot = grid_fig3, width = 4.75, height = 5.25, units = c("in"), dpi=600 )
```


# Boxplot with raw data


```{r}
carbon_plot = ggplot(pcf, aes(x = live.dead, y = bpca.carbon, color = live.dead)) +
  geom_boxplot() +
  geom_point(aes(shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  labs(x=(NULL), y= "g BPCA /\nkg Carbon") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank())+
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

carbon_plot
  
```

```{r}
material_plot = ggplot(pcf, aes(x = live.dead, y = bpca.soil, color = live.dead)) +
  geom_boxplot() +
  geom_point(aes(shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  labs(x=(NULL), y= "g BPCA /\nkg material") +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank())+
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))
material_plot
```
scale_fill_manual(values = alpha(c("blue", "red"), .3))
```{r}
aci_plot = ggplot(pcf, aes(x = live.dead, y = b6.total, color = live.dead)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(color = live.dead, fill = live.dead, shape = bark.char), size = 3, position = position_jitter(w = 0.15, h = 0, seed=10)) +
  labs(x=(NULL), y= "g B6CA /\nTotal BPCA") +
  scale_fill_manual(values = c("tan", "darkgreen")) +
  scale_color_manual(values = c("tan", "darkgreen")) + 
  theme_bw() + 
  theme(legend.position = "none") +  
theme(panel.grid.major = element_blank())+
    theme(axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.line = element_line(colour = "black"))

aci_plot
  
```


```{r fig.height=5.25, fig.width=4.75}
legend = get_legend(legend)
fig2 = cowplot::plot_grid(carbon_plot, material_plot, aci_plot, labels = c("A", "B", "C"), ncol = 1) 
fig2
```



```{r fig.height=5.25, fig.width=4.75}
legend2 = get_legend(legend_bottom)
grid_final2 = ggdraw(fig2) +
  draw_plot(legend2, x =.65, y = .78, width = .3, height = .3) +
    draw_label("*", x = .55, y = .94, size = 36, fontface = "bold") +
  draw_label("*", x = .55, y = .61, size = 36, fontface = "bold")

grid_final2
```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2019-12-19_PyC_live-dead.jpeg", plot = grid_final2, width = 4.75, height = 5.25, units = c("in"), dpi=600 )
```
 
```{r fig.height=5.25, fig.width=4.75}
( grid_final3 = cowplot::plot_grid(fig2, legend, ncol=1, rel_heights = c(1, .06))  )
```
```{r eval=FALSE, include=FALSE}
ggsave("../figures/2019-12-17_PyC_live-dead.jpeg", plot = grid_final3, width = 6, height = 7, units = c("in"), dpi=600 )
```


```{r fig.height=5.25, fig.width=4.75}
legend2 = get_legend(legend_bottom)
grid_sup = ggdraw(fig2) +
  draw_plot(legend2, x =.65, y = .78, width = .3, height = .3) 

grid_sup

```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2019-12-19_PyC_live-dead_sup.jpeg", plot = grid_sup, width = 4.75, height = 5.25, units = c("in"), dpi=600 )
```
**THE END**